`SimSurvey`: an `R` package for comparing the design and analysis of surveys by simulating spatially-correlated populations

### 2020-04-16 analysis/paper/response_letter.Rmd

L326: I did not see how one can use INLA or RandomFields in Appendix S3. This is a similar comment as the one for line 240-242.

*We have added the following in parentheses: see the code behind `sim_ays_covar_sped` [https://github.com/PaulRegular/SimSurvey/blob/master/R/sim_dist_spde.R] for an example of how the `sim_ays_covar` closure was modified to apply a Stochastic Partial Differential Equation approach using the `INLA` package.*

L481-490: the processing time is surprisingly long… it is not an issue in itself but this might refrain some people to using this package… I think some code optimization could help (e.g. vectorize, use matrix operation as much as possible, etc)

*We have stretched our programming skills to the limit to reduce the processing time as much as possible. Of course, said programming skills are largely self-taught, so there is surely room for improvement. While developing the package, we repeatedly profiled our code and vectorized as much as we could, leaning heavily on the `data.table` package. We have also applied parallel processing in places in an attempt to speed up the code. In other parts of the package, we learned that there is little we can do to reduce the processing time (e.g. Cholesky decomposition is a factor limiting the speed of the `sim_distribution` function). Another step to further optimize the code would be to translate some of the core functions to C++, however this is on the wish list and well outside the scope of the project at this point in time. We are hopeful that we have produced enough here to catalyze collaborations that help extend and optimize the package.*

### 2020-04-16 analysis/paper/sim_survey_paper.Rmd
Minor tweaks to rev3 files given feedback from co-authors

Likewise, age frequencies need to be calculated to obtain stratified estimates of abundance at age. This is done by constructing an age-length key, which is the proportion of fish in each length bin that fall into specific ages classes. Once these proportions are calculated, they applied to the bumped up length frequencies, $n_{l,i}$, to approximate age frequencies, $n_{a,i}$, that is:

@@

The equations outlined in the **[Simulate spatial distribution]** section are used in the `make_grid`, `sim_ays_covar` and `sim_parabola` functions, and these functions are used within `sim_distribution` to distribute a population simulated using `sim_abundance` throughout a grid (Table \@ref(tab:distribution)). The output from `make_grid` is a raster object [@hijmans2016] with four layers: `depth`, `cell`, `division` and `strat`. If a more detailed and realistic grid is required, users can manually generate their own survey grid using real data and this grid can be supplied as a raster to `sim_distribution` if the same structure structure and correct projection is used. The package includes a manually constructed survey grid of NAFO Subdivision 3Ps off the southern coast of Newfoundland (named `survey_grid`) and the data-raw folder in the GitHub directory includes the data and code used to construct this grid. However, for simplicity, we use `make_grid` to construct a square grid for a default run of `sim_distribution`. Below we generate and plot (Figure \@ref(fig:grid)) a default grid, another grid with the number of strata increased by increasing the number of `strat_splits`, and another with the number of divisions increased using `n_div` and a linear depth gradient (the sigmoid curve is forced to be linear when `shelf_width` is set to zero).

@@

| **Function call**                   | **Description**                                                             | **Symbol**              |
|:----------------------------------- |:--------------------------------------------------------------------------- | :---------------------- |
| `sim_survey(`                       | &nbsp;                                                                      | &nbsp;                  |
| `+  sim,`                           | Simulated spatial population from `sim_distribution`                        | &nbsp;                  |
| `+  n_sims = 1`                     | Number of times to repeat the survey                                        | &nbsp;                  |
| `+  q = sim_logistic(k = 2,`        | Steepness of logistic curve of catchability                                 | $k$                     |
| `+                   x0 = 3),`      | Midpoint of logistic curve of catchability (age)                            | $x_0$                   |
| `+  trawl_dim = c(1.5, 0.02),`      | Trawl dimensions (distance towed, trawl width; km) - i.e. area trawled      | $A_{\mathrm{trawl}}$    |
| `+  min_sets = 2`                   | Minimum number of sets to conduct per strata                                | &nbsp;                  |
| `+  set_den = 2/1000,`              | Set density (km^-2^)                                                        | $D_{\mathrm{sets}}$     |
| `+  lengths_cap = 500,`             | Maximum number of lengths to collect / set                                  | $M_{\mathrm{lengths}}$  |
| `+  ages_cap = 10,`                 | Maximum number of ages to sample / length group / spatial group ^1^         | $M_{\mathrm{ages}}$     |
| `+  age_sampling = "stratified")`   | Controls whether age sampling is length `"stratified"` or `"random"`        | &nbsp;                  |
| `+  age_length_group = 1)`          | Length group bin size for stratified age sampling (cm)                      | $l_{\mathrm{group}}$    |
| `+  age_space_group = "division")`  | Spatial scale of stratified age sampling (`"division"`, `"strat"`, `"set"`) | $s_{\mathrm{group}}$    |

@@

Results shown in the **[Core functions]** section demonstrates one use of the **`SimSurvey`** package that is focused on evaluating the efficacy of increasing the sampling effort of a stratified random survey of a cod population (see **[S1 Appendix][S1 Appendix: Case study]** for details). These results largely align with expectations from sampling theory that 1) design-based estimators for stratified random surveys are unbiased, and 2) precision is increased by increasing the number of primary sampling units [@Cochran1977]. Specifically, estimates of total abundance and abundance at length are centered around true values and their probability envelopes tighten as set density increases (Figures \@ref(fig:fan1), \@ref(fig:fan2)). Case study results also echo the growing body of literature which concludes that extra sub-sampling is an ineffective means of improving estimates relative to sampling more locations [@Pennington1994; @Pennington2002; @Stewart2014; @Bogstad1995; @Coggins2013; @Zhang2013]. This is exemplified by the relatively large drops in RMSE when set density is increased compared to when sub-sampling effort is increased (Figures \@ref(fig:rank), \@ref(fig:rmse-surface)). Moreover, it appears that it is more advantageous to measure fewer total fish at more locations than measuring many fish at fewer locations (Figure \@ref(fig:rank)). Again, this result was expected because the size-structured spatial clustering imposed by the simulation causes set-to-set variation to be greater than within-set variation in characteristics such as length and age. This structure was imposed because fish caught together tend to have similar characteristics and it is well known that this reduces the effective sample size [@Pennington1994]. Increasing sub-sampling effort, however, should not lead to poorer population estimates like those observed for abundance at age (Figure \@ref(fig:fan3), \@ref(fig:rmse-surface)). Bias, in particular, appears to be a problem with stratified estimates of abundance at age under certain scenarios (Figure \@ref(fig:fan3)); in contrast, there appears to be little bias in the stratified estimates of abundance at length (Figure \@ref(fig:fan2)). This result was unexpected as these estimates of abundance at age were presumed to be unbiased. The contrast between the length and age based analysis indicates that the problem lies with the intervening age-length-key.

Age-length keys have long been used in conjunction with length frequency tables to estimate the age distribution of fish populations over large scales [@fridriksson1934]. By default, `sim_survey` spreads length-stratified age sampling across the division and `run_strat` constructs and applies age-length keys at the division scale. This default was chosen because it is standard protocol for the actual survey the case study is based on. There is, however, a potential cost to the spatial scale of the key. Namely, it is unlikely that one age-length key is representative for the whole region as the probability of being a specific age given length varies in space [@Berg2012]. Because different age groups sometimes occur in different places, the translation of lengths to ages may be biased by the samples used to generate the age-length key. In short, multi-stage sampling of a clustered population violates the assumption that the age-length key is generated from a simple random sample [@Aanes2015]. Bias introduced by this assumption can be resolved by explicitly accounting for the hierarchical sampling by using designed-based estimators [@Aanes2015]. For the case study, a proper design-based estimator will need to account for cluster sampling in a stratified random survey. Following recommendations in Aanes and Vølstad [-@Aanes2015], we used **`SimSurvey`** to 1) conduct a surveys with concurrent length and age sampling at every set, 2) construct and apply age-length keys on a set-by-set basis, and 3) weight the resultant age frequencies at each set using stratified random estimators. This test was accomplished using this code:

@@

This is a minor modification of the code used in the **[`test_surveys`]** section whereby the same simulated population was used (`pop` object) but set density scenarios were reduced to one option, for simplicity, and age sampling protocol and stratified analysis options were changed. That is, 1) `age_length_group` and `age_space_group` were changed to simulate a survey that conducts concurrent length and age sampling at every set (e.g. collect age samples from a maximum of 1 fish per 3 cm length group per set), 2) `alk_scale` was changed to `"set"` to construct and apply age-length keys on a set-by-set basis, and 3) the resultant age frequencies at each set are weighted accordingly using the design-based estimators built into `run_strat`. Again, results are visualized using `plot_age_strat_fan` and `plot_error_surface`:

>>>>

Likewise, age frequencies need to be calculated to obtain stratified estimates of abundance at age. This is done by constructing an age-length key, which is the proportion of fish in each length bin that fall into specific age classes. Once these proportions are calculated, they are applied to the bumped up length frequencies, $n_{l,i}$, to approximate age frequencies, $n_{a,i}$, that is:

@@

The equations outlined in the **[Simulate spatial distribution]** section are used in the `make_grid`, `sim_ays_covar` and `sim_parabola` functions, and these functions are used within `sim_distribution` to distribute a population simulated using `sim_abundance` throughout a grid (Table \@ref(tab:distribution)). The output from `make_grid` is a raster object [@hijmans2016] with four layers: `depth`, `cell`, `division` and `strat`. If a more detailed and realistic grid is required, users can manually generate their own survey grid using real data and this grid can be supplied as a raster to `sim_distribution` if the same structure and correct projection is used. The package includes a manually constructed survey grid of NAFO Subdivision 3Ps off the southern coast of Newfoundland (named `survey_grid`) and the data-raw folder in the GitHub directory includes the data and code used to construct this grid. However, for simplicity, we use `make_grid` to construct a square grid for a default run of `sim_distribution`. Below we generate and plot (Figure \@ref(fig:grid)) a default grid, another grid with the number of strata increased by increasing the number of `strat_splits`, and another with the number of divisions increased using `n_div` and a linear depth gradient (the sigmoid curve is forced to be linear when `shelf_width` is set to zero).

@@

| **Function call**                   | **Description**                                                             | **Symbol**              |
|:----------------------------------- |:--------------------------------------------------------------------------- | :---------------------- |
| `sim_survey(`                       | &nbsp;                                                                      | &nbsp;                  |
| `+  sim,`                           | Simulated spatial population from `sim_distribution`                        | &nbsp;                  |
| `+  n_sims = 1`                     | Number of times to repeat the survey                                        | &nbsp;                  |
| `+  q = sim_logistic(k = 2,`        | Steepness of logistic curve of catchability                                 | $k$                     |
| `+                   x0 = 3),`      | Midpoint of logistic curve of catchability (age)                            | $x_0$                   |
| `+  trawl_dim = c(1.5, 0.02),`      | Trawl dimensions (distance towed, trawl width; km) - i.e. area trawled      | $A_{\mathrm{trawl}}$    |
| `+  min_sets = 2`                   | Minimum number of sets to conduct per strata                                | &nbsp;                  |
| `+  set_den = 2/1000,`              | Set density (km^-2^)                                                        | $D_{\mathrm{sets}}$     |
| `+  lengths_cap = 500,`             | Maximum number of lengths to collect / set                                  | $M_{\mathrm{lengths}}$  |
| `+  ages_cap = 10,`                 | Maximum number of ages to sample / length group / spatial group ^1^         | $M_{\mathrm{ages}}$     |
| `+  age_sampling = "stratified",`   | Controls whether age sampling is length `"stratified"` or `"random"`        | &nbsp;                  |
| `+  age_length_group = 1,`          | Length group bin size for stratified age sampling (cm)                      | $l_{\mathrm{group}}$    |
| `+  age_space_group = "division")`  | Spatial scale of stratified age sampling (`"division"`, `"strat"`, `"set"`) | $s_{\mathrm{group}}$    |

@@

Results shown in the **[Core functions]** section demonstrates one use of the **`SimSurvey`** package that is focused on evaluating the efficacy of increasing the sampling effort of a stratified random survey of a cod population (see **[S1 Appendix][S1 Appendix: Case study]** for details). These results largely align with expectations from sampling theory that 1) design-based estimators for stratified random surveys are unbiased, and 2) precision is increased by increasing the number of primary sampling units [@Cochran1977]. Specifically, estimates of total abundance and abundance at length are centered around true values and their probability envelopes tighten as set density increases (Figures \@ref(fig:fan1), \@ref(fig:fan2)). Our case study results also echo the growing body of literature which concludes that extra sub-sampling is an ineffective means of improving estimates relative to sampling more locations [@Pennington1994; @Pennington2002; @Stewart2014; @Bogstad1995; @Coggins2013; @Zhang2013]. This is exemplified by the relatively large drops in RMSE when set density is increased compared to when sub-sampling effort is increased (Figures \@ref(fig:rank), \@ref(fig:rmse-surface)). Moreover, it appears that it is more advantageous to measure fewer total fish at more locations than measuring many fish at fewer locations (Figure \@ref(fig:rank)). Again, this result was expected because the size-structured spatial clustering imposed by the simulation causes set-to-set variation to be greater than within-set variation in characteristics such as length and age. This structure was imposed because fish caught together tend to have similar characteristics and it is well known that this reduces the effective sample size [@Pennington1994]. Increasing sub-sampling effort, however, should not lead to poorer population estimates like those observed for abundance at age (Figure \@ref(fig:fan3), \@ref(fig:rmse-surface)). Bias, in particular, appears to be a problem with stratified estimates of abundance at age under certain scenarios (Figure \@ref(fig:fan3)); in contrast, there appears to be little bias in the stratified estimates of abundance at length (Figure \@ref(fig:fan2)). This result was unexpected as these estimates of abundance at age were presumed to be unbiased. The contrast between the length and age based analysis indicates that the problem lies with the intervening age-length-key.

Age-length keys have long been used in conjunction with length frequency tables to estimate the age distribution of fish populations over large scales [@fridriksson1934]. By default, `sim_survey` spreads length-stratified age sampling across the division and `run_strat` constructs and applies age-length keys at the division scale. This default was chosen because it is standard protocol for the actual survey the case study is based on. There is, however, a potential cost to the spatial scale of the key. Namely, it is unlikely that one age-length key is representative for the whole region as the probability of being a specific age given length varies in space [@Berg2012]. Because different age groups sometimes occur in different places, the translation of lengths to ages may be biased by the samples used to generate the age-length key. In short, multi-stage sampling of a clustered population violates the assumption that the age-length key is generated from a simple random sample [@Aanes2015]. Bias introduced by this assumption can be resolved by explicitly accounting for the hierarchical sampling by using designed-based estimators [@Aanes2015]. For the case study, a proper design-based estimator will need to account for cluster sampling in a stratified random survey. Following recommendations in Aanes and Vølstad [-@Aanes2015], we used **`SimSurvey`** to 1) conduct a survey with concurrent length and age sampling at every set, 2) construct and apply age-length keys on a set-by-set basis, and 3) weight the resultant age frequencies at each set using stratified random estimators. This test was accomplished using this code:

@@

This is a minor modification of the code used in the **[`test_surveys`]** section whereby the same simulated population was used (`pop` object) but set density scenarios were reduced to one option for simplicity, and age sampling protocol and stratified analysis options were changed. That is, 1) `age_length_group` and `age_space_group` were changed to simulate a survey that conducts concurrent length and age sampling at every set (e.g. collect age samples from a maximum of 1 fish per 3 cm length group per set), 2) `alk_scale` was changed to `"set"` to construct and apply age-length keys on a set-by-set basis, and 3) the resultant age frequencies at each set are weighted accordingly using the design-based estimators built into `run_strat`. Again, results are visualized using `plot_age_strat_fan` and `plot_error_surface`:
